\section{Introduction}\label{sec:intro}

Work-related accidents and diseases kill nearly three million people every year, and an estimated 395 million workers sustain non-fatal occupational injuries over the same period \cite{ilo2023}. The economic burden approaches 4\% of global GDP \cite{ilo2023,lloyds2024}; in the United States alone, work injuries cost \$176.5 billion in 2023 --- more than three billion dollars per week \cite{nsc2023}. In heavy industry, personal protective equipment (PPE) is frequently the last barrier between a hazard and a worker, and safety regulation mandates its use accordingly. A mandate, however, is only as effective as its enforcement. On most industrial sites, enforcement still means a human operator watching a wall of CCTV monitors, expected to notice a missing helmet or an unauthorized entry the moment it happens.

That arrangement fails for a well-documented reason: human vigilance is a rapidly depleting resource. Laboratory studies of sustained attention show detection performance collapsing within 20--35 minutes of continuous monitoring \cite{parasuraman1984,sawin1995}, and reported estimates in the surveillance literature suggest operators can miss up to 95\% of on-screen activity after 22 minutes \cite{velastin2006}. Field evidence is no more reassuring: trained operators monitoring live feeds in diamond-processing plants detected only about half of targeted risk behaviors while generating high false-alarm rates \cite{donald2015}, and attention lapses accumulate steadily over the course of 90-minute monitoring sessions \cite{sandia2014}. The arithmetic of the task makes the problem stark. A single camera produces 525{,}600 minutes of footage per year, and an accident occupies roughly one of them. Year-round manual surveillance pays off only if an operator happens to be fully attentive during that exact minute --- one instant decides the year, and the vigilance literature says the odds are poor.

Automated PPE detection promises to remove this bottleneck, and YOLO-family detectors now dominate the field \cite{ultralytics2023,ppeyolo2024,edgeyolo2024}. Yet published systems are built and evaluated on construction-centric, lab-clean benchmarks; to our knowledge, none has been evaluated on imagery from a real, operating chemical plant. Three gaps produce that vacuum. First, and most fundamentally, the available training data is construction-centric: public benchmarks such as the Roboflow PPE-Detection corpus \cite{roboflowppe}, and even the broader 17-class SH17 dataset \cite{ahmad2024sh17}, concentrate on helmets and high-visibility vests while under-representing both the classes chemical operations depend on, such as gas masks and goggles, and the conditions of a phosphate-processing site: dust, harsh and variable lighting, and site-specific uniforms that generic models mistake for PPE. Second, published pipelines stop at raw detector output. Every false positive lands on a safety officer's desk, and at realistic alert volumes this recreates the vigilance problem one step downstream --- the false-alarm burden documented in manual CCTV monitoring \cite{donald2015} reappears as alert fatigue. Vision-language models capable of reasoning over a flagged frame are only beginning to be explored as a verification layer for detection pipelines \cite{vlmreview2025}. Third, few systems provide a mechanism for adapting to an individual site at all, even though the domain shift between benchmark imagery and a specific plant's cameras is precisely where generic detectors break down. This paper's central experiment measures that breakdown directly.

This paper addresses these gaps with a CCTV-retrofit PPE compliance system developed for, and evaluated at, OCP Group phosphate-processing sites in Morocco. Its core claim is domain-specific rather than architectural: a detector fused and adapted for industrial and chemical conditions --- oblique CCTV viewing angles, gas masks, site-specific uniforms --- outperforms a detector trained only on the standard construction benchmark when both face chemical-plant imagery. Our contributions are:
\begin{itemize}
  \item \textbf{Real-plant central experiment.} A controlled comparison between a generic detector trained solely on the public construction benchmark \cite{roboflowppe} and our industrial-adapted model, evaluated on held-out OCP chemical-plant imagery and on an industrial proxy test set. This comparison --- not a new architecture --- is what the paper claims: evidence that domain adaptation is what makes CV-based PPE monitoring work at an operating plant.
  \item \textbf{Composite domain-adapted dataset.} A fusion of the public PPE-Detection corpus (Roboflow) \cite{roboflowppe}, SH17 \cite{ahmad2024sh17}, and a dedicated gas-mask dataset covering the respirator class public benchmarks omit, together with a self-annotated OCP dataset with weighted sampling favoring site-specific data; this dataset is what the central experiment shows to matter.
  \item \textbf{LLM-based judge stage.} A deployment-oriented verification pass (Qwen2.5-VL \cite{qwen25vl}, served via Ollama) over every flagged detection, designed to suppress false positives arising from occlusion, reflections, distance, and lighting artifacts, and to attach a stated reason to each alert it confirms.
  \item \textbf{Retrofit deployment model.} The pipeline runs on existing camera and network infrastructure; no new hardware is required, and confirmed alerts are pushed to the channels safety staff already use.
\end{itemize}

The remainder of the paper is organized as follows: Section~\ref{sec:related} reviews related work, Section~\ref{sec:methods} details the dataset and system architecture, Section~\ref{sec:results} presents experimental results, Section~\ref{sec:discussion} discusses trade-offs and limitations, and Section~\ref{sec:conclusion} concludes with directions for future work.
