\section{Materials and Methods}
\label{sec:methods}

\subsection{System Overview}
\label{sec:methods-overview}

The system is organized as the four-stage pipeline of Fig.~\ref{fig:pipeline}. A pretrained YOLOv8 detector supplies the perception base (stage 1). It is fine-tuned on a composite, domain-adapted PPE dataset whose final component is self-annotated imagery from OCP site cameras (stage 2). Every detection that indicates a potential violation is re-evaluated by a vision--language judge, which confirms or rejects it and states why (stage 3). Confirmed violations are dispatched as alerts to safety personnel (stage 4).

Two constraints shape this design. The pipeline must retrofit onto CCTV cameras already installed at the sites, which rules out new imaging hardware and favors a small detector (Section~\ref{sec:methods-training}). Its output must also be trustworthy enough that safety staff keep acting on it; a detector that forwards raw alarms merely relocates the vigilance burden it was built to remove, which motivates the verification stage (Section~\ref{sec:methods-judge}). Within this design, stage 2 carries the paper's scientific question: whether a detector trained on generic public benchmarks transfers to an operating chemical plant at all, or whether industrially adapted training data is required. The comparison protocol that answers it is defined in Section~\ref{sec:methods-comparison}.

\begin{figure}[t]
\centering
\resizebox{\columnwidth}{!}{%
\begin{tikzpicture}[
  font=\footnotesize\sffamily,
  >=Latex,
  box/.style={
    draw,
    rounded corners=1.5pt,
    align=center,
    inner sep=3pt,
    minimum height=1.35cm,
    minimum width=2.35cm,
    line width=0.6pt
  },
  stage/.style={box, fill=gray!12},
  arrow/.style={->, thick, shorten >=1pt, shorten <=1pt}
]
  \node[stage] (s1) {%
    \textbf{(1) Detect}\\[1pt]
    YOLOv8n\\[-1pt]
    {\scriptsize on CCTV frames}};
  \node[stage, right=0.55cm of s1] (s2) {%
    \textbf{(2) Adapt}\\[1pt]
    Domain fine-tuning\\[-1pt]
    {\scriptsize fused PPE dataset}};
  \node[stage, right=0.55cm of s2] (s3) {%
    \textbf{(3) Verify}\\[1pt]
    Qwen2.5-VL judge\\[-1pt]
    {\scriptsize confirm / reject}};
  \node[stage, right=0.55cm of s3] (s4) {%
    \textbf{(4) Alert}\\[1pt]
    Dispatch to HSE\\[-1pt]
    {\scriptsize existing infra}};
  \draw[arrow] (s1) -- (s2);
  \draw[arrow] (s2) -- (s3);
  \draw[arrow] (s3) -- (s4);
\end{tikzpicture}%
}
\caption{The four-stage pipeline. (1)~A pretrained YOLOv8 base model provides generic detection capability; (2)~fine-tuning on a composite domain-adapted PPE dataset that includes a self-annotated OCP component adapts it to industrial PPE classes and site conditions; (3)~a Qwen2.5-VL judge re-evaluates each flagged detection and returns a structured verdict with a stated reason; (4)~confirmed violations are dispatched as timestamped alerts to safety personnel over the existing camera and network infrastructure.}
\label{fig:pipeline}
\end{figure}

\subsection{Dataset Construction}
\label{sec:methods-dataset}

No single public corpus covers the class mix of a chemical-processing site, so the training set fuses three public sources and one site-specific corpus under a single class map (Fig.~\ref{fig:datasets}). Roboflow PPE-Detection \cite{roboflowppe} contributes a helmet- and vest-centred detection benchmark typical of construction imagery. SH17 \cite{ahmad2024sh17} contributes 17 classes of PPE items and human body parts across varied industrial scenes. A dedicated respirator dataset supplies instances of the gas-mask class that construction-skewed benchmarks barely cover.
% TODO before submission: cite the gas-mask dataset source here once it is chosen
% for the rerun (add the entry to references.bib and \cite it).
All sources are merged into a unified ten-class map: Person, helmet, vest, gloves, boots, goggles, no\_helmet, no\_gloves, no\_goggle, and gas mask. The explicit negative classes (no\_helmet, no\_gloves, no\_goggle) let the detector localize the \emph{absence} of an item directly instead of inferring it from a missing detection.

\begin{figure}[t]
\centering
\resizebox{\columnwidth}{!}{%
\begin{tikzpicture}[
  font=\footnotesize\sffamily,
  >=Latex,
  box/.style={
    draw,
    rounded corners=1.5pt,
    align=center,
    inner sep=3pt,
    minimum height=0.9cm,
    minimum width=3.4cm,
    line width=0.6pt
  },
  src/.style={box, fill=gray!12},
  proc/.style={box, fill=gray!25, minimum height=1.35cm, minimum width=2.7cm},
  arrow/.style={->, thick, shorten >=1pt, shorten <=1pt},
  note/.style={font=\scriptsize\sffamily, fill=white, inner sep=1pt}
]
  \node[src] (d1) {Roboflow PPE-Detection\\[-1pt]{\scriptsize public $\cdot$ helmets, vests}};
  \node[src, below=0.3cm of d1] (d2) {SH17\\[-1pt]{\scriptsize public $\cdot$ 17 industrial classes}};
  \node[src, below=0.3cm of d2] (d3) {Gas-mask dataset\\[-1pt]{\scriptsize public $\cdot$ respirators}};
  \node[src, below=0.3cm of d3] (d4) {OCP site cameras\\[-1pt]{\scriptsize self-annotated}};

  \coordinate (mid) at ($(d2.east)!0.5!(d3.east)$);
  \node[proc, right=1.9cm of mid] (fuse) {\textbf{Fusion}\\[1pt]explicit class-ID remap\\[-1pt]{\scriptsize unified 10-class map}};
  \node[proc, right=0.8cm of fuse] (model) {\textbf{YOLOv8n}\\[1pt]fine-tuning\\[-1pt]{\scriptsize $640\times640$ $\rightarrow$ \texttt{best.pt}}};

  \draw[arrow] (d1.east) -- node[note, pos=0.45, above, sloped] {ID remap} ($(fuse.west)+(0,0.45)$);
  \draw[arrow] (d2.east) -- node[note, pos=0.45, above, sloped] {ID remap} ($(fuse.west)+(0,0.15)$);
  \draw[arrow] (d3.east) -- node[note, pos=0.45, below, sloped] {oversampled} ($(fuse.west)+(0,-0.15)$);
  \draw[arrow] (d4.east) -- node[note, pos=0.45, below, sloped] {sampling weight $\uparrow$} ($(fuse.west)+(0,-0.45)$);
  \draw[arrow] (fuse) -- (model);
\end{tikzpicture}%
}
\caption{Dataset composition feeding the detector fine-tune (stage 2 of Fig.~\ref{fig:pipeline}). Four sources are merged under a unified ten-class map with explicit label-ID remapping before any image enters training; SH17 classes outside the PPE scope are pruned. The gas-mask source is oversampled to counter class imbalance, and the self-annotated OCP corpus receives a higher sampling weight so that batches skew toward deployment-realistic imagery.}
\label{fig:datasets}
\end{figure}

Fusing annotation sets is not a file-concatenation exercise, and the remapping step deserves emphasis because its failure mode is silent. Each source ships label files whose integer class IDs index that source's own ontology: SH17's IDs 0--16 reference its 17-class list, not the fused ten-class map. Under naive merging, IDs that happen to fall inside the fused range are reinterpreted as the wrong class, while every image containing a label ID at or above the fused class count is discarded by the training framework as corrupt, with no error raised. In a preliminary naive merge, the second mode alone silently rejected roughly 43\% of the fused training images. Our fusion pipeline therefore performs explicit ID remapping before merging: each SH17 class is either mapped to its fused-map ID or deliberately excluded (e.g., body-part classes outside the PPE scope), the mapping is recorded, and all label files are rewritten against the unified map before any image enters training.

The gas-mask class receives targeted treatment. Respiratory protection is among the most safety-relevant PPE items at chemical-processing sites, yet it is under-represented in public data, which skews toward construction. The fused map reserves a gas-mask class, populated primarily by the dedicated respirator source; because gas-mask instances remain scarce relative to helmets and vests, they are further reinforced through oversampling and the augmentation policy of Section~\ref{sec:methods-preprocessing}.

The fourth dataset component is a self-annotated corpus from OCP site cameras, capturing conditions public data lacks: phosphate dust, harsh lighting transitions, and site-specific uniforms and equipment. This corpus has been collected and annotated against the fused class map using Roboflow's annotation tooling, and is included in the training mix. OCP samples receive a higher sampling weight than public samples, so each batch is biased toward deployment-realistic imagery while the public data preserves visual diversity. A held-out portion of the OCP corpus, disjoint from all training data, forms the chemical-plant evaluation set of the central comparison (Section~\ref{sec:methods-comparison}); publication of individual site frames remains subject to imagery clearance. The fused dataset is partitioned into \phX\% training, \phX\% validation, and \phX\% test splits, stratified across classes.

\subsection{Preprocessing}
\label{sec:methods-preprocessing}

All images are resized to $640\times640$ pixels. The augmentation policy targets the visual failure modes of industrial CCTV. HSV jitter perturbs hue, saturation, and value to simulate the colour and contrast shifts produced by dust haze and glare. Horizontal flipping doubles pose variety. Random erasing imitates partial occlusion by equipment and other workers. Mosaic augmentation composites four training images to vary object scale and context, and is disabled for the final training epochs so that late optimization sees realistically composed frames.

\subsection{Model Training}
\label{sec:methods-training}

The detector is YOLOv8n \cite{ultralytics2023}, the nano variant of YOLOv8: 3.0M parameters, 8.1 GFLOPs at $640\times640$, and 6.3 MB of weights. The small footprint is a deliberate choice rather than a compromise, since retrofit deployment means inference must run on commodity hardware alongside existing video infrastructure. Training runs on Google Colab for \phX{} epochs at $640\times640$ with batch size 16. Optimizer selection is left to the framework's automatic policy, which configures an SGD-family optimizer with initial learning rate 0.01, momentum 0.937, and weight decay $5\times10^{-4}$; the first three epochs apply learning-rate warmup. The best checkpoint by validation fitness is exported as \texttt{best.pt} and used in all subsequent stages. Detection accuracy is reported in Section~\ref{sec:results}.

\subsection{Generic-vs-Adapted Comparison Protocol}
\label{sec:methods-comparison}

The central experiment asks whether generic benchmark training transfers to an operating chemical plant. The \emph{generic baseline} is the detector trained solely on the Roboflow PPE-Detection corpus \cite{roboflowppe}: 308 images over nine construction-oriented classes, reporting 58.1\% precision, 64.4\% recall, and 59.1\% mAP@50 on its own validation split --- competitive numbers for the benchmark it was built on, and representative of what a practitioner deploying a public benchmark model obtains off the shelf. The \emph{industrial-adapted model} is the fused-dataset detector of Section~\ref{sec:methods-training}.

Both models are evaluated on two test sets that neither saw during training. The first is the held-out OCP site imagery of Section~\ref{sec:methods-dataset} --- the headline comparison, since it is exactly the deployment domain: oblique CCTV angles, phosphate dust, gas masks, and site uniforms. The second is an \emph{industrial proxy} set assembled from held-out SH17 industrial scenes and respirator images, which approximates chemical-plant content while remaining publishable without imagery clearance. Because the two models carry different class inventories, comparison metrics are computed over the classes common to both maps (person, helmet, vest, gloves, and their negative states), at identical input resolution and confidence thresholds. The baseline differs from our model in training-set size as well as domain, a confound discussed in Section~\ref{sec:discussion}; the comparison measures what the public benchmark model actually delivers on plant imagery, which is the deployment-relevant question.

\subsection{LLM Judge Stage}
\label{sec:methods-judge}

A detector that forwards every flagged frame directly to a human recreates the problem it is meant to solve. Donald et al.\ found that trained CCTV operators in diamond-processing plants detected only about half of targeted risk behaviours while producing a high false-alarm rate \cite{donald2015}; an automated detector with an unfiltered alarm stream shifts that burden downstream, where accumulating false alerts teach staff to ignore the system. A verification stage is therefore necessary by design, independent of any particular error rate measured for our detector. Vision--language verification of detector output is an emerging pattern in the literature \cite{vlmreview2025}, and we adopt it here.

Stage three re-evaluates each flagged detection with Qwen2.5-VL \cite{qwen25vl}, a vision--language model served through Ollama.\footnote{\url{https://ollama.com}} The pilot uses a cloud-hosted instance; the same model is self-hostable on-premises, in which case no frame ever leaves the site network. The judge receives the cropped detection region, the full frame for context, and detection metadata (flagged class, detector confidence, camera ID). It returns a structured JSON verdict, one of \texttt{confirm}, \texttt{reject}, or \texttt{recheck}, together with a confidence value and a one-sentence reason. Only \texttt{confirm} reaches a human; \texttt{recheck} re-queues the detection for evaluation on subsequent frames. The judge prompt is:

\begin{quote}\small
You are a safety-compliance verifier for industrial CCTV. A PPE detector flagged the attached image region as ``\textless class\textgreater'' with confidence \textless conf\textgreater; the full frame is provided for context. Decide whether this is a genuine violation. Reason about: occlusion by equipment or other workers, reflections and glare, distance and image resolution, lighting and dust conditions, and whether pose or clothing could mimic a violation. Reply in JSON: \{``verdict'': ``confirm''\,\textbar\,``reject''\,\textbar\,``recheck'', ``confidence'': 0--1, ``reason'': ``\textless one sentence\textgreater''\}.
\end{quote}

Two properties distinguish this stage from simpler false-positive filters. First, raising the detector's confidence threshold trades recall for precision blindly: it discards low-confidence true violations together with the artifacts. The judge instead reasons about \emph{why} a detection may be wrong (occlusion, glare, distance) and can preserve a low-confidence detection that is nevertheless genuine. Second, the reason field accompanies every alert, giving safety staff an explainable and auditable basis for each intervention rather than an opaque score.

We present the judge as an architectural contribution and evaluate it qualitatively in this paper (Section~\ref{sec:results}); we do not claim a measured false-positive reduction. Its inclusion is motivated by the human-factors evidence above rather than by our own measurements, and quantitative benchmarking of the judge is identified as future work.

\subsection{Alert and Deployment Architecture}
\label{sec:methods-alerts}

When the judge confirms a violation, the dispatch stage pushes an alert to safety personnel containing the camera ID, the violation class, the judge's stated reason, and a timestamped clip of the event. Alerts travel over the sites' existing camera and network infrastructure, and detector inference runs on commodity compute, so the pipeline requires no new capital hardware. Coverage is continuous across all connected feeds: unlike a human operator, the pipeline has no attention span to exhaust.
