\section{Discussion}\label{sec:discussion}

A detector that flags a missing helmet is not yet a safety system. Whether the pipeline described here improves safety at a phosphate site depends on two surrounding design decisions: adapting the detector to the environment it must watch, and verifying its outputs before they reach a human. This section examines both, then states the limits of the present evaluation.

\subsection{Why Domain Adaptation Matters}

Public PPE corpora are dominated by construction imagery: near-frontal views, daylight, helmets and high-visibility vests \cite{roboflowppe,ppeyolo2024}. A phosphate processing site violates each of these assumptions. Airborne dust lowers contrast and softens object boundaries; fixed CCTV cameras observe workers from high, oblique angles at distances curated datasets rarely contain; and the equipment itself differs, with gas masks and goggles mattering as much as helmets. The composite dataset is a direct response to these mismatches. Fusing the Roboflow PPE-Detection corpus with SH17 \cite{ahmad2024sh17} broadens pose, scale, and class coverage beyond any single source, and the gas-mask augmentation targets the class most specific to chemical processing. On the fused public data the detector reaches \phX{} mAP@50 at \phX{}\,ms per frame, which supports the retrofit claim: usable accuracy at real-time speed on commodity hardware. The remaining gap is deliberately assigned to the third stage of the data design, weighted fine-tuning on self-annotated OCP imagery, which is in progress and not evaluated here. The comparison it enables, generic versus site-weighted on identical OCP frames, is the experiment the dataset was built for.

\subsection{Trade-offs of the LLM Judge Stage}

The takeaway of this paper is that detection alone does not produce a safety system; it relocates the vigilance problem. An unverified detector converts a monitoring burden into an alert-triage burden: safety staff receive a stream in which false positives dominate, habituate to it, and eventually miss the true events, the same decrement documented for CCTV operators \cite{donald2015,sawin1995}. Verification is what makes an alert worth acting on, which is why the judge sits inside the pipeline rather than beside it.

The stage has real costs. Each flagged event incurs a vision-language-model inference measured in seconds rather than milliseconds; because verification runs off the per-frame path this does not affect detection throughput, but it does delay notification. More serious is the asymmetry of judge errors: a rejected true violation is costlier than the false alert it would have replaced, so over-suppression, not latency, is the risk that most needs measurement. Serving location is a third trade-off. Ollama-hosted Qwen2.5-VL \cite{qwen25vl} can run on-premises, keeping frames inside the site network, or in the cloud, which removes the need for an on-site GPU at the price of egress and added latency. The pilot used cloud hosting; on-premises serving is the intended production configuration given the sensitivity of site imagery.

We do not claim a measured false-positive reduction for the judge; the evidence in this paper is architectural and qualitative. The measurement path is nonetheless concrete: assemble a labeled alert set of detector-flagged events, have site safety staff mark each as a genuine violation or a false alarm, run the judge over the set, and report its confirm/reject confusion matrix against that ground truth, with the judge's false-reject rate as the safety-critical headline. This protocol is the first item of future work.

\subsection{Limitations}

\begin{itemize}
  \item \textbf{Small and distant objects:} gloves and goggles occupy few pixels in wide CCTV views, and the nano-scale detector loses accuracy on them well before helmets or vests are affected.
  \item \textbf{Dust and weather:} heavy dust, rain, and low light plausibly push accuracy below the reported figures; robustness under these conditions has not been measured.
  \item \textbf{Gas-mask scarcity:} augmentation raises the sample count but not the diversity of the class, which remains under-represented relative to its importance in chemical processing.
  \item \textbf{Judge evaluation:} qualitative only. No false-positive-reduction figure has been measured, and none is claimed.
  \item \textbf{Camera placement:} the system inherits the blind spots of the existing CCTV layout; better detection cannot compensate for uncovered areas.
  \item \textbf{OCP fine-tuning:} designed and in progress; all site-specific performance claims await its results.
\end{itemize}
