\section{Conclusion and Future Work}\label{sec:conclusion}

Industrial PPE monitoring fails for lack of sustained attention, not for lack of cameras. We presented a pipeline that retrofits detection and verification onto the CCTV infrastructure of an operating chemical plant: a YOLOv8 detector fine-tuned on a composite, industrially adapted dataset \cite{roboflowppe,ahmad2024sh17,roboflowgasmasks}, and a vision-language-model judge that reviews each flagged event before it becomes an alert. The central result is the generic-versus-adapted comparison: a detector trained only on the standard construction benchmark reaches \phX{} mAP@50 on held-out OCP chemical-plant imagery, against \phX{} for the adapted model. The gap is evidence that site adaptation, not architectural novelty, is what makes CV-based PPE monitoring work at an operating plant; the adapted detector delivers this at \phX{}\,ms per frame on commodity hardware.

Future work proceeds in order of what the argument needs most. First, a size-matched control for the central comparison, to separate the effect of domain coverage from training-set volume. Second, quantitative benchmarking of the judge on a labeled alert set, with a confirm/reject confusion matrix against human ground truth. Third, a full-corpus re-evaluation of SH17's validation and test splits, to check whether the 300-image compute-motivated subsample (Section~\ref{sec:methods-dataset}) shifts reported metrics relative to the full source, and a class-aware replacement for the current uniform random cap. Fourth, an ablation restoring gas-mask oversampling to confirm that omitting it, in favor of budget headroom elsewhere in the pipeline, did not measurably cost gas-mask recall. Fifth, expansion of the OCP dataset across additional sites and seasons to further reduce residual domain shift. Sixth, temporal tracking and person re-identification, so that compliance is assessed per worker over time rather than per frame. Seventh, edge deployment to cut per-camera latency and network load \cite{edgeyolo2024}, and integration with access-control systems so that verified violations can gate entry to hazardous zones. The ordering is deliberate: the first four items harden the paper's own claims, and cost less than any extension that follows.

\section*{Acknowledgment}
% TODO: confirm wording with partners
The authors thank OCP Group for site access and collaboration on the deployment study, and EMSI Marrakech for institutional support.

\section*{Data Availability}
The public datasets used in this work are available from their original sources \cite{roboflowppe,ahmad2024sh17,roboflowgasmasks}. Imagery collected at OCP sites is proprietary and cannot be made publicly available.
